---
title: "Associations Between 11 Forms of Discipline and Child Outcomes Across 60 Countries"
author:
  - name: Kaitlin P. Ward
    affiliation: University of Michigan
    affiliation_url: https://www.umich.edu
  - name: Andrew Grogan-Kaylor
    url: https://agrogan1.github.io/
    affiliation: University of Michigan
    affiliation_url: https://www.umich.edu
  - name: Shawna J. Lee
    affiliation: University of Michigan
    affiliation_url: https://www.umich.edu  
  - name: Julie Ma
    affiliation: University of Michigan
    affiliation_url: https://www.umich.edu  
  - name: Garrett T. Pace 
    affiliation: University of Michigan
    affiliation_url: https://www.umich.edu
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    highlight: haddock
    toc: yes
citation_url: https://agrogan1.github.io/
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```


```{r}

library(maps) # maps

library(maptools) # maptools

library(sp) # spatial data

library(countrycode) # manipulate country names and codes

library(dplyr) # data wrangling

library(tidyr) # tidy data

library(sf) # simple features

library(ggplot2) # beautiful graphs

library(plotly) # interactive graphs

library(tibble) # data frames

```

```{r}

data(wrld_simpl) # world map data

global_data <- st_as_sf(wrld_simpl) %>% # make an sf dataset
  select(-POP2005) # remove POP2005

# get MICS countries

country <- c("Afghanistan", "Algeria",  "Argentina",  
             "Bangladesh",  "Barbados",  "Belarus",  
             "Belize",  "Benin",  "Bosnia and Herzegovina",  
             "Cameroon",  "Central African Republic",  "Chad",  
             "Democratic Republic of the Congo",  
             "Republic of the Congo",  
             "Costa Rica",  "Cote d'Ivoire",  "Dominican Republic",  
             "El Salvador",  "Eswatini",  "Ghana",  
             "Guinea",  "Guinea Bissau",  "Guyana",  
             "Iraq",  "Jamaica",  "Kazakhstan",  
             "Kenya",  "Kosovo",  "Kyrgyzstan",  
             "Laos", "Macedonia",  "Madagascar",  
             "Malawi",  "Mali",  "Mauritania",  
             "Mexico",  "Moldova",  "Mongolia",  
             "Montenegro",  "Nepal",  "Nigeria",  
             "Pakistan",  "Panama",  "Paraguay",  
             "Sao Tome and Principe",  "Senegal",  "Serbia",  
             "Sierra Leone",  "Somalia",  "St. Lucia",  
             "State of Palestine",  "Suriname",  "Thailand",  
             "The Gambia",  "Togo",  "Trinidad and Tobago",  
             "Tunisia",  "Turkmenistan",  "Ukraine",  
             "Uruguay",  "Vietnam",  "Zimbabwe")

# convert to ISO3

country_iso <- countrycode(country, 
                           'country.name', 
                           'iso3c')

# MICS is an sf object that is subset of global_data

MICS <- global_data %>% 
  filter(ISO3 %in% country_iso)
  
```

# Objective

To test associations between 11 parental discipline behaviors and child aggression, distraction, and prosocial peer relations across low- and middle-income countries (LMICs).

# Study Design 

Data came from the fourth (2009-2013) and fifth (2012-2017) rounds of the UNICEF Multiple Indicator Cluster Surveys (MICS). Survey responses publicly available as of July 2020 were included. Data were restricted to household respondents with focal children under 5 years. The final analytic sample included 163,345 respondents across 60 LMICs. Data were analyzed using Bayesian multilevel logistic regression.

```{r, fig.cap="Locations of MICS Countries", layout = "l-page"}

library(leaflet)

leaflet() %>%
  setView(0, 0, zoom = 1.5) %>%
  addProviderTiles(providers$Stamen.Terrain) %>%
  addPolygons(data = MICS, 
              fillOpacity = .75, 
              color = "#5b92e5", 
              label = MICS$NAME,
              highlightOptions = highlightOptions(color = "red", 
                                                  weight = 2,
                                                  bringToFront = TRUE))

```

# Results 

## Main Effects

Verbal reasoning (80%) and shouting (66%) were the most common parental discipline behaviors toward young children. Psychological and physical aggression were associated with higher child aggression and distraction. Verbal reasoning was associated with lower aggression (OR = 0.92, 95% CI [0.86, 0.99]) and higher prosocial peer relations (OR = 1.30, 95% CI [1.20, 1.42]).

Taking away privileges was associated with higher distraction (OR = 1.09, 95% CI [1.03, 1.15] and lower prosocial peer relations (OR = 0.92, 95% CI [0.87, 0.98]). Giving the child something else to do was associated with higher distraction (OR = 1.06, 95% CI [1.01, 1.12].

## Random Slopes

Random slopes suggested statistically credible cultural variation in the associations between parenting behaviors and child socio-emotional outcomes.

# Data Visualization

```{r, fig.cap="Association of Discipline with Child Outcomes", layout = "l-page"}

myORs <- tribble(
  ~Discipline, ~Aggression, ~Distraction, ~Prosocial,
  "Shouted at child", 1.2, 1.08, NA, 
  "Called child names", 1.29, 1.14, 0.88, 
  "Shook child", 1.2, 1.08, 0.82, 
  "Spanked child", 1.12, NA, NA, 
  "Hit/slapped child on arm/leg", 1.12, 1.07, NA, 
  "Hit child with object", 1.17, NA, NA, 
  "Hit/slapped child on face/head/ears", 1.24, 1.06, 0.86, 
  "Beat as hard as one could", NA, NA, NA, 
  "Took privileges away from child", NA, 1.09, 0.92, 
  "Explained why behavior was wrong", 0.92, NA, 1.3, 
  "Gave child something else to do", NA, 1.06, NA)

p1 <- myORs %>%
  pivot_longer(cols = 2:4,
               names_to = "outcome") %>%
  mutate(change = ifelse(value > 1, "increases", "decreases")) %>%
  ggplot(aes(x = Discipline,
             y = value,
             fill = outcome,
             text = paste0("\"",
                          Discipline, 
                          "\" ",
                          change,
                          " ",
                          tolower(outcome)))) +
  # geom_col() +
  geom_point(size = 5) +
  geom_hline(yintercept = 1.0, color = "red") +
  ylim(.7, 1.5) +
  coord_flip() +
  facet_wrap(~outcome) +
  scale_fill_viridis_d() +
  theme_minimal()

ggplotly(p1, tooltip = "text")

```

# Conclusions

Psychological and physical aggression were disadvantageous for children's socio-emotional development across countries.

Only verbal reasoning was associated with positive child socio-emotional development.

Greater emphasis should be dedicated to reducing parental use of psychological and physical aggression across cultural contexts.

