---
title: "Interesting and Noteworthy Findings From MICS Analyses"
date: "`r Sys.Date()`"
format: 
  html:
    theme: 
      light: united
      dark: vapor
    toc: false
    number-sections: true
    anchor-sections: true
    highlight-style: haddock
    code-fold: true
    execute:
      warning: false
    comments:
      hypothesis: true
  pdf:
    toc: false
    toc-depth: 3
    number-sections: true
    colorlinks: true
    highlight-style: haddock
    execute:
      echo: false
      warning: false
    shift-heading-level-by: 0  
  docx:
    toc: true
    number-sections: true
    highlight-style: haddock
    execute:
      echo: false
      warning: false     
title-block-banner: "#009edb"
subtitle: "Implications For The Cross Cultural Study Of Parenting And Child Development"
bibliography: ../MICS.bib
biblio-style: apalike
nocite: |
  @Grogan-Kaylor2021, @WardA, @WardC, @ward_grogan-kaylor_ma_pace_lee_2021, @Pace2019, @Ma2022
citation_url: https://agrogan1.github.io/research/MICS
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.margin=TRUE)

```

```{r}

# call libraries

library(ggplot2)

library(ggthemes)

library(sf) # simple (spatial) features

library(fontawesome) # fontawesome from Rstudio

```

```{r}

# get data

global_data <- read_sf("../../mapping/shapefiles/wrld_simpl/wrld_simpl.shp")

MICS <- read_sf("../../mapping/shapefiles/MICS/MICS.shp")

```

# Data 

The findings below are drawn from studies using the *Multiple Indicator Cluster Surveys* (MICS) conducted by UNICEF. Data include more than 200,000 families from a diverse sample of approximately 60 Low and Middle Income (LMICS) around the world.

```{r, echo=FALSE, fig.cap="Countries in MICS"}
#| label: fig-MICS
#| column: margin

ggplot(global_data) +
  geom_sf(fill = "gray93",
          size = .25) +
  geom_sf(data = MICS, 
          aes(fill = continent)) +
  scale_fill_tableau(name = "Continent",
                     palette = "Tableau 10", 
                     type = "regular") +
  # scale_fill_viridis_d(option = "turbo") +
  labs(title = "Countries in MICS") +
  # theme_minimal() +
  theme_void()

```

# No Spanking

Spanking is associated with decreases in child socio-emotional development & with increases in child aggression.

```{r}
#| column: margin

fa("child", fill = "darkblue", width = "3em")

```

# No Insults

Name-calling & psychological aggression may have effects that are as harmful as those of physical punishment.

```{r}
#| column: margin

fa("child", fill = "darkblue", width = "3em")

```

# No Physical Or Psychological Punishments

Though common, these parenting behaviors are consistently associated with decreases in child socio-emotional development & with increases in child aggression.

```{r}
#| column: margin

fa("child", fill = "darkblue", width = "3em")

```


```{r, echo=FALSE, results='asis'}

# print a reference header when not HTML5

if(knitr::opts_knit$get("rmarkdown.pandoc.to") != "html5") {
  
  cat("# References")
  
}

```
