---
title: "Parenting and Child Development"
date: "today"
format: 
  html:
    theme: 
      light: cosmo
      dark: vapor
    toc: true
    number-sections: false
    anchor-sections: true
    highlight-style: haddock
    code-fold: true
    execute:
      warning: false
    comments:
      hypothesis: false
  docx:
    toc: false
    number-sections: false
    highlight-style: haddock
    reference-doc: markstat.docx
    execute:
      echo: false
      warning: false  
  PrettyPDF-pdf:
    keep-tex: true         
subtitle: "Implications For The Cross Cultural Study Of Parenting And Child Development from MICS Data"
bibliography: ../MICS.bib
biblio-style: apalike
nocite: |
  @Grogan-Kaylor2021, @WardA, @WardC, @ward_grogan-kaylor_ma_pace_lee_2021, @Ward2023, @Pace2019, @Ma2022
citation_url: https://agrogan1.github.io/research/MICS
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.margin=TRUE)

```

```{r}

# call libraries

library(ggplot2)

library(ggthemes)

library(sf) # simple (spatial) features

library(fontawesome) # fontawesome from Rstudio

```

```{r}

# get data

global_data <- read_sf("../../mapping/shapefiles/wrld_simpl/wrld_simpl.shp")

MICS <- read_sf("../../mapping/shapefiles/MICS/MICS.shp")

```


```{r dev.args = list(bg = 'transparent')}
#| label: fig-MICS
#| echo: false
#| fig-cap: "Countries in MICS"
#| fig-pos: "H"
#| fig-height: 3

# NB: dev.args in chunk arguments
# sets chunk to have transparent background
# need to also included transparency in ggplot theme()

ggplot(global_data) +
  geom_sf(fill = "gray93",
          size = .25) +
  geom_sf(data = MICS, 
          aes(fill = continent)) +
  scale_fill_viridis_d(name = "Region") +
  # scale_fill_tableau(name = "Region",
  #                    palette = "Tableau 10", 
  #                    type = "regular") +
  # scale_fill_viridis_d(option = "turbo") +
  labs(title = "Countries in MICS") +
  # theme_minimal() +
  theme_void() +
  theme(panel.background = element_rect(fill = "transparent", 
                                        colour = NA),
        plot.background = element_rect(fill = "transparent", 
                                       colour = NA))

```

# Spanking Hurts Children

Spanking is associated with decreased child socio-emotional development & increased child aggression.

# Insults Contribute to Harm

Name-calling & psychological aggression may have effects that are as harmful as those of physical punishment.

# Physical and Psychological Punishment Slow Child Development

Harsh punishment is consistently associated with decreased child socio-emotional development & more child aggression.

# Eliminating Spanking Reduces Abuse

Simulations with MICS data suggest that eliminating spanking would result in a large reduction in global rates of physical child abuse.

# Use Positive Discipline!

Positive discipline–in the form of verbal reasoning–is associated with improved child well-being across MICS countries.

# Positive discipline benefits children.

```{r, echo=FALSE, results='asis'}

# print a reference header when not HTML5

if(knitr::opts_knit$get("rmarkdown.pandoc.to") != "html5") {
  
  cat("# References")
  
}

```




