---
title: "Work With Performance Measurement Initiative"
author:
  - name: Andy Grogan-Kaylor 
    url: https://agrogan1.github.io/
    affiliation: University of Michigan
    affiliation_url: https://www.umich.edu
date: "`r Sys.Date()`"
format: 
  html:
    theme: 
      light: united
      dark: vapor
    toc: true
    number-sections: true
    anchor-sections: true
    highlight-style: haddock
    code-fold: true
    execute:
      warning: false
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

# Collaborating with the William Davidson Institute {.smaller}

I am interested in interventions that may reduce poverty, and thereby contribute to child and family wellbeing.  Thus, I collaborate with the [Performance Measurement Initiative](http://wdi.umich.edu/initiatives/performance-measurement/) at the  [William Davidson Institute](http://wdi.umich.edu/) at the Ross School of Business. 

```{r, eval=TRUE}

knitr::include_url("https://wdi.umich.edu/news/qa-faculty-affiliate-looks-for-real-world-applications/",
                   height = "400px")

```

# Work With Vision Spring

Our first published article from our work, *Connecting Poverty to Purchase in Informal Markets* can be accessed [here](http://doi.org/10.1002/sej.1173).

# Work With Danone Ecosystem

A direct link is [here](https://wdi.umich.edu/wp-content/uploads/Semilla-Report_04-29-2016_Extract.pdf). 

```{r}

knitr::include_url("https://wdi.umich.edu/wp-content/uploads/Semilla-Report_04-29-2016_Extract.pdf",
                   height = "400px")

```

# Program Evaluation Workshops

Several years ago, one of our major activities was to conduct a series of several day impact assessment workshops that teach members of non-profits and socially conscious businesses how to conduct outcome evaluations.  We conducted workshops in Ann Arbor, New York City, and Johannesburg, South Africa. 

# Business and Culture Course

Currently, we are engaged in an evaluation of the [Business and Culture](https://businessandculture.org/about/) course, a partnership between the Ross School of Business and several institutions in Egypt, Lebanon and Libya.

```{r}

knitr::include_url("https://businessandculture.org/",
                   height = "400px")

```









