---
title: "Performance Measurement Initiative"
description: |
  Collaborating with the William Davidson Institute.
author:
  - name: Andy Grogan-Kaylor 
    url: https://agrogan1.github.io/
    affiliation: University of Michigan
    affiliation_url: https://www.umich.edu
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    highlight: haddock
    toc: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

```{r, eval=FALSE, echo=FALSE}

library(plotly)

plot_geo() %>%
  layout(geo = list(showland = TRUE,  
                    landcolor = toRGB("grey"),
                    showcountries = TRUE,
                    showocean=FALSE, 
                    projection = list(type = 'orthographic'),
                    lataxis = list(showgrid = TRUE,
                                   gridcolor = toRGB("black")),
                    lonaxis = list(showgrid = TRUE,
                                   gridcolor = toRGB("black")))) 

```


I am interested in interventions that may reduce poverty, and thereby contribute to child and family wellbeing.  Thus, I collaborate with the [Performance Measurement Initiative](http://wdi.umich.edu/initiatives/performance-measurement/) at the  [William Davidson Institute](http://wdi.umich.edu/) at the Ross School of Business. See [here](http://wdi.umich.edu/programs-projects/danone-ecosystem-fund-semilla-impact-assessment/) for one of our recent projects.
 
Several years ago, one of our major activities was been to conduct a series of several day impact assessment workshops that teach members of non-profits and socially conscious businesses how to conduct outcome evaluations.  We conducted workshops in Ann Arbor, New York City, and Johannesburg, South Africa. 
 
Our first published article from our work, “Connecting Poverty to Purchase in Informal Markets” can be accessed [here](http://doi.org/10.1002/sej.1173).

Currently, we are engaged in an evaluation of the [Business and Culture](https://businessandculture.org/about/) course, a partnership between the Ross School of Business and several institutions in Egypt, Lebanon and Libya.

```{r}

knitr::include_url("https://businessandculture.org/",
                   height = "600px")

```









