---
title: "Country Level Bans on Corporal Punishment"
author: "Andy Grogan-Kaylor" 
date: "today"
format:
  html:
    toc: true
    number-sections: true
    anchor-sections: true
    theme:
      light: yeti
      dark: vapor
    lightbox: true
    css: styles.css
---


```{r call_libraries}
#| eval: true
#| echo: false
#| output: false
#| warning: false

library(readr) # get data

library(foreign) # read and write other formats

library(maps) # basic mapping

library(countrycode) # code country names

# library(leaflet) # web maps

library(ggplot2) # nice graphs

library(ggthemes) # nice themes for ggplot2

library(ggrepel) # distance between ggplot labels

library(dplyr) # data wrangling

library(tibble) # library for updated dataframes

library(DT) # nifty data table

library(plotly) # interactive plots

library(fontawesome) # Font Awesome icons

# library(xaringanExtra)
# 
# xaringanExtra::use_panelset()

```

```{r get_data}
#| eval: true
#| echo: false

cpbans <- tribble( 
  ~type, ~year.of.prohibition, ~country, ~country_code, ~total.number.of.bans,
  "CP Ban", 2024, "Tajikistan", "TJK", 67,
  "CP Ban", 2024, "Lao PDR", "LAO", 66,
  "CP Ban", 2022, "Mauritius", "MUS", 65,
  "CP Ban", 2022, "Zambia", "ZMB", 64, 
  "CP Ban", 2021, "Colombia", "COL", 63, 
  "CP Ban", 2021, "Korea", "KOR", 62, 
  "CP Ban", 2021, "Guinea", "GIN", 61, 
  "CP Ban", 2020, "Seychelles", "SYC", 60, 
  "CP Ban", 2019, "Japan", "JPN", 59, 
  "CP Ban", 2019, "Georgia", "GEO", 58, 
  "CP Ban", 2019, "South Africa", "ZAF", 57, 
  "CP Ban", 2019, "France", "FRA", 56, 
  "CP Ban", 2019, "Kosovo", "RKS", 55, 
  "CP Ban", 2018, "Nepal", "NPL", 54, 
  "CP Ban", 2016, "Montenegro", "MNE", 53, 
  "CP Ban", 2017, "Lithuania", "LTU", 52, 
  "CP Ban", 2016, "Slovenia", "SVN", 51, 
  "CP Ban", 2016, "Paraguay", "PRY", 50, 
  "CP Ban", 2016, "Mongolia", "MNG", 49, 
  "CP Ban", 2015, "Peru", "PER", 48, 
  "CP Ban", 2015, "Ireland", "IRL", 47, 
  "CP Ban", 2015, "Benin", "BEN", 46, 
  "CP Ban", 2015, "Andorra", "AND", 45, 
  "CP Ban", 2014, "Nicaragua", "NIC", 44, 
  "CP Ban", 2014, "Estonia", "EST", 43, 
  "CP Ban", 2014, "San Marino", "SMR", 42, 
  "CP Ban", 2014, "Argentina", "ARG", 41, 
  "CP Ban", 2014, "Cape Verde", "CPV", 40, 
  "CP Ban", 2014, "Bolivia", "BOL", 39, 
  "CP Ban", 2014, "Brazil", "BRA", 38, 
  "CP Ban", 2014, "Malta", "MLT", 37, 
  "CP Ban", 2013, "Honduras", "HND", 36, 
  "CP Ban", 2013, "Macedonia", "MKD", 35, 
  "CP Ban", 2011, "South Sudan", "SSD", 34, 
  "CP Ban", 2010, "Albania", "ALB", 33, 
  "CP Ban", 2010, "Congo", "COG", 32, 
  "CP Ban", 2010, "Kenya", "KEN", 31, 
  "CP Ban", 2010, "Tunisia", "TUN", 30, 
  "CP Ban", 2010, "Poland", "POL", 29, 
  "CP Ban", 2008, "Liechtenstein", "LIE", 28, 
  "CP Ban", 2008, "Luxembourg", "LUX", 27, 
  "CP Ban", 2008, "Moldova", "MDA", 26, 
  "CP Ban", 2008, "Costa Rica", "CRI", 25, 
  "CP Ban", 2007, "Togo", "TGO", 24, 
  "CP Ban", 2007, "Spain", "ESP", 23, 
  "CP Ban", 2007, "Venezuela", "VEN", 22, 
  "CP Ban", 2007, "Uruguay", "URY", 21, 
  "CP Ban", 2007, "Portugal", "PRT", 20, 
  "CP Ban", 2007, "New Zealand", "NZL", 19, 
  "CP Ban", 2007, "Netherlands", "NLD", 18, 
  "CP Ban", 2006, "Greece", "GRC", 17, 
  "CP Ban", 2005, "Hungary", "HUN", 16, 
  "CP Ban", 2004, "Romania", "ROU", 15, 
  "CP Ban", 2004, "Ukraine", "UKR", 14, 
  "CP Ban", 2003, "Iceland", "ISL", 13, 
  "CP Ban", 2002, "Turkmenistan", "TKM", 12, 
  "CP Ban", 2000, "Germany", "DEU", 11, 
  "CP Ban", 2000, "Israel", "ISR", 10, 
  "CP Ban", 2000, "Bulgaria", "BGR", 9, 
  "CP Ban", 1999, "Croatia", "HRV", 8, 
  "CP Ban", 1998, "Latvia", "LVA", 7, 
  "CP Ban", 1997, "Denmark", "DNK", 6, 
  "CP Ban", 1994, "Cyprus", "CYP", 5, 
  "CP Ban", 1989, "Austria", "AUT", 4, 
  "CP Ban", 1987, "Norway", "NOR", 3, 
  "CP Ban", 1983, "Finland", "FIN", 2, 
  "CP Ban", 1979, "Sweden", "SWE", 1, 
)

```

```{r wrangle_data}
#| eval: true
#| echo: false
#| warning: false

save(cpbans, file = "CPBans.RData") # save to R

write.dta(cpbans, "CPBans.dta") # write to Stata

write.csv(cpbans, file = "CPBans.csv") # save to CSV

# make new data file of ALL Countries

AllCountries <- read_csv("AllCountries.csv") # read data

CPBans_w_AllCountries <- merge(cpbans, 
                               AllCountries, 
                               by = "country_code",
                               all = TRUE) # merge


CPBans_w_AllCountries$continent <- 
  countrycode(CPBans_w_AllCountries$country_code, 
              origin = 'iso3c',
              destination = 'continent') 

save(CPBans_w_AllCountries, 
     file = "CPBans_w_AllCountries.RData") # save to R

write.dta(CPBans_w_AllCountries, 
          "CPBans_w_AllCountries.dta") # write to Stata

```

# Background 

**`r max(cpbans$total.number.of.bans)`** countries have instituted country wide bans upon the use of corporal punishment with children.  The most recent country to institute a ban is **`r cpbans$country[1]`**.

Information on countries that have banned corporal punishment available from  [endcorporalpunishment.org](http://endcorporalpunishment.org/)

# Visualizations

::: {.panel-tabset}

## Interactive Globe

```{r make_globe}
#| eval: true
#| echo: false

plot_geo(cpbans) %>% 
  add_trace(locations = ~country_code, 
            color = ~year.of.prohibition,
            z = ~year.of.prohibition, 
            text = ~paste(country, 
                          "<br>year of prohibition:", 
                          year.of.prohibition),
            hoverinfo = "text",
            marker = list(size = 10)) %>%
  layout(title = "Countries With Bans on Corporal Punishment of Children", 
         geo = list(showland = FALSE,
                    showcountries = TRUE,
                    projection = list(type = 'orthographic',
                                      rotation = list(lon = -30,
                                                      lat = 10,
                                                      roll = 0)
                                      ))) %>%
  colorbar(title = 'year of prohibition') 

```


## Map

```{r make_map}
#| eval: true
#| echo: false
#| fig-height: 6

plot_geo(cpbans) %>%
  add_trace(locations = ~country_code, 
            color = ~year.of.prohibition,
            z = ~year.of.prohibition,
            hoverinfo = "text",
            text = ~paste(country, 
                          "<br>year of prohibition:", 
                          year.of.prohibition)) %>%
  layout(title = "Countries With Bans on Corporal Punishment of Children",
         geo = list(showland = FALSE,
                    showcountries = TRUE,
                    projection = list(type = 'natural earth'))) %>%
  colorbar(title = 'year of prohibition')

```

## Motion Chart

```{r make_motion_chart_data}
#| eval: true
#| echo: false
#| warning: false

# create data with repeated observations for animation

cpbans$number.repetitions <- 
  max(cpbans$year.of.prohibition) + 
  1 - 
  cpbans$year.of.prohibition

cpbans2 <- data.frame(rep(cpbans$country, 
                          cpbans$number.repetitions),
                      rep(cpbans$year.of.prohibition, 
                          cpbans$number.repetitions),
                      rep(cpbans$total.number.of.bans, 
                          cpbans$number.repetitions))


names(cpbans2)[1] <- "country"

names(cpbans2)[2] <- "year.of.prohibition"

names(cpbans2)[3] <- "number_bans"

cpbans2 <- cpbans2 %>% 
  group_by(country) %>% 
  mutate(id = n() - row_number()) %>%
  mutate(year = year.of.prohibition + id)

cpbans2$country_iso <- countrycode(cpbans2$country, 
                                 'country.name', 
                                 'iso3c')

```

```{r plotly_motion_chart}
#| eval: true
#| echo: false
#| fig-height: 8
#| warning: false

# motion chart

cpbans2 %>% 
  arrange(year, number_bans) %>% 
  plot_ly(x = ~year.of.prohibition, 
        y = ~number_bans,
        ids = ~country,
        frame = ~year, 
        color = ~year.of.prohibition, 
        type = 'scatter',
        mode = 'markers',
        text = ~country,
        textfont = list(color = toRGB("red")),
        textposition = 'middle right',
        marker = list(size = 10),
        showlegend = FALSE) %>%
  hide_colorbar() %>%
  layout(title = 'Country Level Bans on Corporal Punishment', 
         xaxis = list(title = "year", 
                      range = c(1975, 2025)),
         yaxis = list(title = "total number of bans")) 

```

## Data Table

```{r make_data_table}
#| eval: true
#| echo: false

mynewdata <- subset(cpbans, 
                    select = c(total.number.of.bans, 
                               country, 
                               year.of.prohibition))

datatable(mynewdata, 
          rownames = FALSE,
          # filter = 'top',
          caption = 'Countries With Bans on Corporal Punishment of Children',
          colnames = c('Number' = 1, 
                       'Country' = 2, 
                       'Year of Prohibition' = 3), 
          extensions = 'Buttons', 
          options = list(
            dom = 'Bfrtip',
            buttons = c('copy', 
                        'csv', 
                        'excel', 
                        'pdf', 
                        'print'))
          )

```

## Graph

```{r make_graph}
#| eval: true
#| echo: false
#| fig-height: 6
#| warning: false
#| message: false

p1 <- ggplot(cpbans, 
             aes(x = year.of.prohibition, 
                 y = total.number.of.bans)) + 
  geom_line(stat = "smooth", 
            method = "loess",
            color = "grey",
            span = 0.5,
            linewidth = 3,
            alpha = .25) +
  geom_point(aes(color = total.number.of.bans),
             # color = "blue", 
             size = 1) +
  geom_text(aes(x = year.of.prohibition - .25 ,
                label = country),
            hjust = 1, 
            size = 2) +
  ggtitle("Total Number of Corporal Punishment Bans by Year") +
  xlim(1975, 
       max(cpbans$year.of.prohibition)) +
  xlab("year") + 
  ylab("total number of bans") +
  theme_minimal() +
  scale_color_viridis_c(name = "Total Number \nof Bans")

p1

```


```{r make_interactive_graph}
#| eval: false
#| echo: false
#| fig-height: 8
#| warning: false

plot_ly(data = cpbans, 
        x = ~year.of.prohibition, 
        y = ~total.number.of.bans,
        color = ~year.of.prohibition, 
        type = 'scatter',
        mode = 'marker',
        marker = list(size = 5),
        text = ~paste(country, 
                      '<br>', 
                      'year of ban:',
                      year.of.prohibition)) %>%
  add_annotations(x = cpbans$year.of.prohibition,
                  y = cpbans$total.number.of.bans,
                  text = cpbans$country,
                  font = list(size = 8), 
                  showarrow = FALSE, 
                  xanchor = 'right') %>%
  layout(title = 'Country Level Bans on Corporal Punishment', 
         xaxis = list(title = "year"),
         yaxis = list(title = "total number of bans")) %>% 
  rangeslider()

```

:::




