---
title: "Meta-Analysis of Research on Corporal Punishment"
author:
  - name: Andy Grogan-Kaylor 
    url: https://agrogan1.github.io/
    affiliation: University of Michigan
    affiliation_url: https://www.umich.edu/
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

```{r, layout="l-screen", fig.height=4}

library(readxl)

CP_Meta_Summary_Results <- read_excel("data/CP Meta Summary Results.xlsx")

library(ggplot2)

library(plotly)

# p1_plotly <- CP_Meta_Summary_Results %>% 
#   plot_ly(x = ~d,
#           y = ~Outcome,
#           color = ~d,
#           text = ~paste(Outcome, "<br>Cohen's d:", d), 
#           hoverinfo = 'text',
#           orientation = 'h',
#           type = "bar") 
# 
# p1_plotly

p1 <- ggplot(CP_Meta_Summary_Results,
             aes(x = Outcome,
                 y = d,
                 # label = Outcome,
                 fill = d)) + 
  geom_bar(stat = "identity") +
  scale_fill_viridis_c() +
  coord_flip() +
  labs(title= "Effect Sizes of Physical Punishment",
       x = "Outcome",
       y = "Cohen's d Effect Size",
       caption = "Gershoff and Grogan-Kaylor, 2016") +
  theme(legend.position = "none",
        title = element_text(size = rel(.5)),
        axis.text = element_text(size = rel(.75)))

ggplotly(p1, tooltip="fill")

```

**Gershoff, E. T., & Grogan-Kaylor, A. (2016). Spanking and Child Outcomes: Old Controversies and New Meta-Analyses. Journal of Family Psychology. 30(4), 453-469. http://dx.doi.org/10.1037/fam0000191**

> This article received a [considerable amount of media attention](media-mentions-CP-meta.html).

Whether spanking is helpful or harmful to children continues to be the source of considerable debateamong both researchers and the public. This article addresses 2 persistent issues, namely whether effect sizes for spanking are distinct from those for physical abuse, and whether effect sizes for spanking arerobust to study design differences. Meta-analyses focused specifically on spanking were conducted on atotal of 111 unique effect sizes representing 160,927 children. Thirteen of 17 mean effect sizes weresignificantly different from zero and all indicated a link between spanking and increased risk fordetrimental child outcomes. Effect sizes did not substantially differ between spanking and physical abuse or by study design characteristics.




